Intersection Types, Sub-formula Property, and the Functional Character of the Lambda Calculus

### 2018-08-09 DenotSem/determ.tex
spell check

A lot of progress can happen in six years, so I sent an email to
Olivier. He replied,
\begin{quote}
``Indeed! I now have two different sequent-calculus systems which are
  equivalent to BCD subtyping and satisfy the sub-formula property.  I
  am currently writing a paper on this but it is not ready yet.''
\end{quote}
and he attached the paper draft and the Coq mechanization. What great
timing!  Furthermore, Olivier would be presenting the paper, titled
\emph{Intersection Subtyping with Constructors}, at the Workshop on
Intersection Types and Related System in Oxford on July 8, part of the
Federated Logic Conference (FLOC). I was planning to attend FLOC
anyways, for the DOMAINS workshop to celabrate Dana Scott's 85th
birthday.

@@

\item Case $C'=C_1\sqcap C_2$:\\
  We already know that $C'$ is an atom, so we have a contradiction
  and this case is vacously true.
\end{itemize}

@@

We shall represent variables with de Bruijn indices, so an environment
$\Gamma$ is a sequence of types.  The denotational semantics of the CBV
$\lambda$-calculus is defined as follows.
\begin{align*}
  \ESEM{n}\Gamma &= \UP{n} \\
  \ESEM{x}\Gamma &=  \IF x < |\Gamma| \THEN \UP{\Gamma[k]} \ELSE \emptyset \\
  \ESEM{\lambda e }\Gamma &= \{ A \mid \WF{A} \text{ and } \FSEM{A}e\Gamma \} \\
  \ESEM{e_1\app e_2}\Gamma &= \left\{ C\, \middle| \exists A,B.\;
      \begin{array}{l}
      A \in \ESEM{e_1}\Gamma,
      B \in \ESEM{e_2}\Gamma,\\
      A <: B \to C, \text{ and } \WF{C}
      \end{array}
        \right\} \\
  \ESEM{f(e_1,e_2)}\Gamma &=
      \left\{ C\, \middle| \begin{array}{l}
       \exists A,B,n_1,n_2.\; A \in \ESEM{e_1}\Gamma, B \in \ESEM{e_2}\Gamma,\\
      \qquad A <: n_1, B <: n_2, \SEM{f}(n_1,n_2) <: C, \WF{C} 
      \end{array} \right\} \\
  \ESEM{\IF e_1 \THEN e_2 \ELSE e_3}\Gamma &=
    \left\{ B\, \middle| \exists A.\;
    \begin{array}{l}A \in \ESEM{e_1}\Gamma, A <: n,\\
           n = 0 \Rightarrow B \in \ESEM{e_3}\Gamma,\\
           n \neq 0 \Rightarrow B \in \ESEM{e_2}\Gamma
    \end{array}
    \right\}
    \\[2ex]
   \FSEM{n}e\Gamma &= \mathit{false} \\
   \FSEM{A \sqcap B}e \Gamma &= \FSEM{A}e\Gamma \text{ and } \FSEM{B}e\Gamma\\
   \FSEM{A \to B}e \Gamma &= B \in \ESEM{e} (A, \Gamma)
\end{align*}

@@

%%  LocalWords:  denotationally ctx subexpression eq denot lrcl env
%%  LocalWords:  iff Hah pdf subtyping contra distributivity BCD et
%%  LocalWords:  Barendregt al Fundamenta Informaticae sequent

>>>>

A lot of progress can happen in six years, so I sent an email to
Olivier. He replied,
\begin{quote}
``Indeed! I now have two different sequent-calculus systems which are
  equivalent to BCD subtyping and satisfy the sub-formula property.  I
  am currently writing a paper on this but it is not ready yet.''
\end{quote}
and he attached the paper draft and the Coq mechanization. What great
timing!  Furthermore, Olivier would be presenting the paper, titled
\emph{Intersection Subtyping with Constructors}, at the Workshop on
Intersection Types and Related System in Oxford on July 8, part of the
Federated Logic Conference (FLOC). I was planning to attend FLOC
anyways, for the DOMAINS workshop to celebrate Dana Scott's 85th
birthday.

@@

\item Case $C'=C_1\sqcap C_2$:\\
  We already know that $C'$ is an atom, so we have a contradiction
  and this case is vacuously true.
\end{itemize}

@@

We shall represent variables with de Bruijn indices, so an environment
$\Gamma$ is a sequence of types.  The denotational semantics of the
call-by-value $\lambda$-calculus is defined as follows.
\begin{align*}
  \ESEM{n}\Gamma &= \UP{n} \\
  \ESEM{x}\Gamma &=  \IF x < |\Gamma| \THEN \UP{\Gamma[k]} \ELSE \emptyset \\
  \ESEM{\lambda e }\Gamma &= \{ A \mid \WF{A} \text{ and } \FSEM{A}e\Gamma \} \\
  \ESEM{e_1\app e_2}\Gamma &= \left\{ C\, \middle| \exists A,B.\;
      \begin{array}{l}
      A \in \ESEM{e_1}\Gamma,
      B \in \ESEM{e_2}\Gamma,\\
      A <: B \to C, \text{ and } \WF{C}
      \end{array}
        \right\} \\
  \ESEM{f(e_1,e_2)}\Gamma &=
      \left\{ C\, \middle| \begin{array}{l}
       \exists A,B,n_1,n_2.\; A \in \ESEM{e_1}\Gamma, B \in \ESEM{e_2}\Gamma,\\
      \qquad A <: n_1, B <: n_2, \SEM{f}(n_1,n_2) <: C, \WF{C} 
      \end{array} \right\} \\
  \ESEM{\IF e_1 \THEN e_2 \ELSE e_3}\Gamma &=
    \left\{ B\, \middle| \exists A.\;
    \begin{array}{l}A \in \ESEM{e_1}\Gamma, A <: n,\\
           n = 0 \Rightarrow B \in \ESEM{e_3}\Gamma,\\
           n \neq 0 \Rightarrow B \in \ESEM{e_2}\Gamma
    \end{array}
    \right\}
    \\[2ex]
   \FSEM{n}e\Gamma &= \mathit{false} \\
   \FSEM{A \sqcap B}e \Gamma &= \FSEM{A}e\Gamma \text{ and } \FSEM{B}e\Gamma\\
   \FSEM{A \to B}e \Gamma &= B \in \ESEM{e} (A, \Gamma)
\end{align*}

@@

%%  LocalWords:  denotationally ctx subexpression eq denot lrcl env
%%  LocalWords:  iff Hah pdf subtyping contra distributivity BCD et
%%  LocalWords:  Barendregt al Fundamenta Informaticae sequent Coq th
%%  LocalWords:  FLOC Scott's Olivier's disjunction reflexivity dom
%%  LocalWords:  subtype Denotational denotational de Bruijn indices
%%  LocalWords:  Subsumption inlining
